---
title: "Problem Set #4"
author: "Margaret Sun"
date: "2/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Install relevant packages

library(gt)
library(tidyverse)
library(janitor)
library(readxl)
library(lubridate)


orig <- read_csv("ps_4_elections-poll-nc09-3.csv",
                 col_types = cols(
                   .default = col_character(),
                   turnout_scale = col_double(),
                   turnout_score = col_double(),
                   w_LV = col_double(),
                   w_RV = col_double(),
                   final_weight = col_double(),
                   timestamp = col_datetime(format = ""))) %>% 
  # Tidy data
  clean_names() %>% 
  filter(!is.na(response),
         !is.na(race_eth),
         !is.na(final_weight),
         educ != "[DO NOT READ] Refused")
  
```


## Question 1
```{r Question 1a, echo=FALSE}

# Filter data for Democratic candidates, counts Democratic candidates

Q1a <- orig %>% filter(response == "Dem") %>% nrow()
```

There were `r Q1a` respondents who supported the Democratic candidate.

```{r Question 1b, echo=FALSE}

# Counts Republican and Undecided responses, finds difference between Republican and Undecided responses

Q1b <- orig %>% 
  filter(response %in% c("Rep", "Und")) %>% 
  group_by(response) %>% 
  count()

Q1b_ans <- Q1b$n[1] - Q1b$n[2]
```

There were `r Q1b_ans` more respondents who favored the Republican candidate than who were Undecided.

```{r Question 1c, echo=FALSE}

# Select gender and gender_combined, filter for where values are different

Q1c <- orig %>% 
  select(gender, gender_combined) %>% 
  filter(gender != gender_combined) %>% 
  count()
```

There are two gender variables (gender and gender_combined). There are `r Q1c` individuals for whom these variables have different values.

```{r Question 1d, echo=FALSE}

# Filters for value "White" under race_eth but not file_race_black

Q1d <- orig %>% 
  filter(race_eth == "White") %>% 
  filter(file_race_black != "White") %>% 
  count()
```

There are `r Q1d` respondents listed as “White” under race_eth who are not listed as “White” under file_race_black.

```{r Question 1e, echo=FALSE}

# Filters for first response time of Rep

rep_first <- orig %>% 
  filter(response == "Rep") %>% 
  arrange(timestamp) %>% 
  slice(1) %>% 
  select(timestamp)

# Filters for first response time of Dem

dem_first <- orig %>% 
  filter(response == "Dem") %>% 
  arrange(timestamp) %>% 
  slice(1) %>% 
  select(timestamp)

# Calculates difference rounded to the nearest minute between first response of Rep and first response of Dem

Q1e <- round(as.numeric(rep_first - dem_first))
```

The first response of Rep came `r Q1e` minutes (rounded to the nearest minute) before the first response of Dem.


## Question 2
```{r make_chart, echo=FALSE, results="asis"}

Q2 <- orig %>% 
  
  # Select relevant columns
  select(response, race_eth, final_weight) %>% 
  group_by(race_eth, response) %>% 
  summarize(total = sum(final_weight, na.rm = TRUE)) %>%   
  filter(race_eth != "[DO NOT READ] Don't know/Refused") %>% 
  spread(key =  response, value = total) %>% 
  mutate(all = Dem + Rep + Und + `3`) %>% 
  mutate(Dem = Dem / all) %>% 
  mutate(Rep = Rep / all) %>% 
  mutate(Und = Und / all) %>% 
  select(-all) %>% 
  
  # One of the biggest pieces of black magic incantation in R is ungroup(). (I
  # did not mention this in class.) Summary: Whenever you group a tibble (as we
  # do above) the grouping stays with an resulting object, until you explicitly
  # ungroup() it. That can't ever hurt things (right? TFs?) and it often helps,
  # as in this case.
  
  ungroup() %>% 
  
  # You will have a chance to explore many other gt commands in problem set #4.
  # I added two extras that we did not get to in class.
  
  gt() %>% 
    tab_header(
      title = "Polling Results in Arizon 6th Congressional District") %>% 
    
    cols_label(
      race_eth = "",
      Dem = "DEM.",
      Rep = "REP.",
      Und = "UND."
      ) %>%
  
    fmt_percent(columns = vars(Dem, Rep, Und),
                decimals = 0) %>% 
  
    # This little pipe is that incantation to take this pretty table, turn it
    # into html, and send it to the md file we are creating. Future versions of
    # gt will probably have a better way of doing this. Indeed, does anyone know
    # of one?

    as_raw_html() %>% as.character() %>% cat()
  

```


