---
title: "Problem Set #4"
author: "Margaret Sun"
date: "2/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Install relevant packages

library(gt)
library(tidyverse)
library(janitor)
library(readxl)
library(lubridate)
library(ggplot2)

orig <- read_csv("ps_4_elections-poll-nc09-3.csv",
                 col_types = cols(
                   .default = col_character(),
                   turnout_scale = col_double(),
                   turnout_score = col_double(),
                   w_LV = col_double(),
                   w_RV = col_double(),
                   final_weight = col_double(),
                   timestamp = col_datetime(format = ""))) %>% 
  # Tidy data
  
  clean_names() %>% 
  filter(!is.na(response),
         !is.na(race_eth),
         !is.na(final_weight),
         educ != "[DO NOT READ] Refused")
  
```


## Question 1
```{r Question 1a, echo=FALSE}

# Filter data for Democratic candidates, counts Democratic candidates

Q1a <- orig %>% filter(response == "Dem") %>% nrow()
```

There were `r Q1a` respondents who supported the Democratic candidate.

```{r Question 1b, echo=FALSE}

# Counts Republican and Undecided responses, finds difference between Republican and Undecided responses

Q1b <- orig %>% 
  filter(response %in% c("Rep", "Und")) %>% 
  group_by(response) %>% 
  count()

Q1b_ans <- Q1b$n[1] - Q1b$n[2]
```

There were `r Q1b_ans` more respondents who favored the Republican candidate than who were Undecided.

```{r Question 1c, echo=FALSE}

# Select gender and gender_combined, filter for where values are different

Q1c <- orig %>% 
  select(gender, gender_combined) %>% 
  filter(gender != gender_combined) %>% 
  count()
```

There are two gender variables (gender and gender_combined). There are `r Q1c` individuals for whom these variables have different values.

```{r Question 1d, echo=FALSE}

# Filters for value "White" under race_eth but not file_race_black

Q1d <- orig %>% 
  filter(race_eth == "White") %>% 
  filter(file_race_black != "White") %>% 
  count()
```

There are `r Q1d` respondents listed as “White” under race_eth who are not listed as “White” under file_race_black.

```{r Question 1e, echo=FALSE}

# Filters for first response time of Rep

rep_first <- orig %>% 
  filter(response == "Rep") %>% 
  arrange(timestamp) %>% 
  slice(1) %>% 
  select(timestamp)

# Filters for first response time of Dem

dem_first <- orig %>% 
  filter(response == "Dem") %>% 
  arrange(timestamp) %>% 
  slice(1) %>% 
  select(timestamp)

# Calculates difference rounded to the nearest minute between first response of Rep and first response of Dem

Q1e <- round(as.numeric(rep_first - dem_first))
```

The first response of Rep came `r Q1e` minutes (rounded to the nearest minute) before the first response of Dem.


## Question 2
```{r make_chart, echo=FALSE, results="asis"}

Q2 <- orig %>% 
  
  # Select relevant columns
  
  select(response, race_eth, final_weight) %>% 
  group_by(race_eth, response) %>% 
  
  # Finds total weights for each race_eth and response combination
  
  summarize(total = sum(final_weight, na.rm = TRUE)) %>%   
  
  # Rearranges data into target table groupings
  
  spread(key =  response, value = total, fill = 0) %>% 
  
  # Account for no Asian responses to Und column with na.rm = TRUE
  
  mutate(all = Dem + Rep + Und + `3`, na.rm = TRUE) %>% 
  
  # Calculates responses as percentages in target table
  
  mutate(Dem = Dem / all, 
         Rep = Rep / all,
         Und = Und / all) %>% 
  select(-`3`, -all, -na.rm) %>% 
  
  ungroup() %>% 
  
  # Reorder variables according to target table
  
  slice(match(c("White", "Black", "Hispanic", "Asian", "Other"), race_eth))
  
 
 # Make table in gt
  
  gt(Q2) %>% 

    # Provides title and source  
    
    tab_header(title = "Polling Results of 9th Congressional District North Carolina") %>%
    tab_source_note(source_note = "Source: New York Times Upshot/Siena College 2018 Live Polls") %>% 
    
    # Labels columns of table
    
    cols_label(
      race_eth = "",
      Dem = "DEM.",
      Rep = "REP.",
      Und = "UND."
      ) %>%

    
    # Convert all values to percentages
    
    fmt_percent(columns = vars(Dem, Rep, Und), decimals = 0) %>% 
    
    # Replace row 4 NA with hyphen to reflect target table
  
    fmt_missing(columns = vars(Und), rows = 4) %>% 
    
    # How to print the sample size and motivation for organizing companies in column 1
      
    # Color-coding for cells in table
    
    tab_style(style = cells_styles(bkgd_color = "#fbfafb", text_color = "#666666"), locations = cells_data()) %>%
    tab_style(style = cells_styles(bkgd_color = "#ffffff", text_color = "#666666"), locations = cells_data(columns = 1)) %>%
    tab_style(style = cells_styles(bkgd_color = "#d0202d", text_color = "#ffffff"), locations = cells_data(columns = 3, rows = 1)) %>% 
  tab_style(style = cells_styles(bkgd_color = "#d0202d", text_color = "#ffffff"), locations = cells_data(columns = 3, rows = 3)) %>%
  tab_style(style = cells_styles(bkgd_color = "#d0202d", text_color = "#ffffff"), locations = cells_data(columns = 3, rows = 4)) %>% 
  tab_style(style = cells_styles(bkgd_color = "#1f87c5", text_color = "#ffffff"), locations = cells_data(columns = 2, rows = 2)) %>% 
  tab_style(style = cells_styles(bkgd_color = "#1f87c5", text_color = "#ffffff"), locations = cells_data(columns = 2, rows = 5)) %>% 

  

    as_raw_html() %>% as.character() %>% cat()
  

```

## Question 3

```{r violin_plot, echo = FALSE}

ggplot(orig, aes(x = educ, y = final_weight)) +
  
  # Creates violin plot
  
  geom_violin() +
  
  geom_jitter(alpha = 0.5, width = 0.21, size = 0.8) +
  
  coord_flip() +
  
  scale_x_discrete(limits=c("Grade school", "High school", "Some college or trade school", "Bachelors' degree", "Graduate or Professional Degree")) +
  
  xlab(NULL) +
  
  ylab("Weight Given to Respondent in Calculating Poll Results") +
  
  labs(title = "More Educated Matter Less in North Carolina 9th",
       subtitle = "Poll gives more weight to people who are less likely to participate in polls",
       caption = "New York Times Upshot/Siena College 2018 live polls")
```


```{r, echo=FALSE, include=FALSE}

# Read in ps_3_enrollment.xlsx spreadsheet
# Clean data, set resulting dataframe as "x"

x <- read_xlsx("ps_3_enrollment.xlsx", skip = 3) %>% 
  janitor::clean_names() %>% 
  
  # Filter for existing courses
  
  filter(!is.na(course_section_code)) %>% 
  
  # Filter for courses with more than 5 undergraduates 
  
  filter(u_grad > 5) %>% 
  
  # Filter for courses with a majority of undergraduates
  
  filter(u_grad / total > 0.5)

# Print number of courses

nrow(x)


# Arranges undergraduate enrollment in descending order
# Print first 9 rows, finds course with 9th highest undergraduate enrollment

x %>% 
  arrange(desc(u_grad)) %>% 
  head(x, n=9)
  


# Finds course with 9th highest undergraduate enrollment

x %>% 
  
  # Groups data by department
  
  group_by(course_department) %>% 
  
  # Finds total undergraduate enrollment in each department
  summarise(total = sum(u_grad)) %>% 
  
  # Arranges total undergraduate enrollment in each department in descending order
  arrange(desc(total)) %>% 
  
  # Prints head (first 6 values), finds department with 6th highest total undergraduate enrollment
  head()



# Finds total undergraduate enrollment in Economics department 

x %>% 
  filter(course_department == "Economics") %>% 
  summarise(total = sum(u_grad)) 

```




## Colleagues

Simone Chu, Igor Morzan